Molecular dynamics simulation of synchronization of a driven particle

### 2021-04-12 manuscript/manuscript_ajp_version3.tex
rejiggered the introduction following, main topic, define, illustrate, transition

\section{Introduction} 
Synchronization is a universal phenomena
in which
individual oscillators change frequency due
to external stimuli.\cite{Pikovsky2003}
The
flickering patterns of
candle flames 
mediated by temperature fluctuations,\cite{Okamoto2016}
the vibrations of singing wine glasses interacting 
through sound waves,\cite{Arane2009}
and metronomes vibrating through a supporting platform\cite{Jia2015}
are examples of in-phase coupled oscillations. 
Biological systems benefit from cooperative
synchronization --
birds  coordinate their wing flaps
to optimize energy use during flight,\cite{Portugal2014}
frogs alternate croaking patterns,\cite{Aihara2014}
humans clap in time with music,\cite{Tranchant2016}
and 
at a cellular level, 
neurons simultaneously fire in cardiac muscle\cite{MartinHall1999}
and brain tissue.\cite{Singer1999}
%DM ended paragraph with fewer examples, added transition sentence
Mechanisms of synchronization
can be WHAT

Synchronized phase-locking or mode-locking 
first appeared in the scientific literature
in 1665 with 
Huygens'  experiments on the
motions of synchronized 
pendula in wall-mounted clocks.\cite{Bennett2002}
A locked-mode is an integer frequency ratio.
In Huygens clocks,
the pendula were observed to swing at the same rate
in the same direction (1:1 mode) or
opposite directions ($-1$:1 mode).
A 2:1 mode occurs when a simple pendulum four
times the length of another swings with twice the period.
%DM transition sentence
Simulations and laboratory experiments
of mode-locked synchronization
can help provide insights to a diverse range of 
oscillation phenomena.

%DM better main topic
Experiments with colloids

Complex dynamics such as synchronized mode-locking
can be studied with
colloid particles.
In experiments %DM moved the ``in experiments'' to second sentence
typical colloids are   
plastic spheres suspended in
de-ionized water or silica beads suspended in organic solvent.
Because 
colloids are large 
and move slowly, the
particle positions 
can be measured in real time 
with an optical camera.\cite{Pertsinidis2001}
Experimental measurements of the
step-by-step dynamics of 
colloids performing phase-locked motions
are useful for
understanding synchronization
at a single particle level.\cite{Juniper2015,Juniper2017}

Light is a
tool for manipulating the colloidal environment
to alter synchronization patterns.
Colloids can be trapped
with 
radiation pressure from 
a laser beam.\cite{Ashkin1997}
A colloid centered in an optical trap is 
uniformly bombarded by photons. 
Off-center colloids 
experience a net force
due to uneven photon collisions across
the particle surface.
Depending on the 
location of the particle in the trap,
the radiation pressure either moves colloids toward the center 
or ejects it from the trap.
Diffraction gratings can create
more complex light environments, 
such as periodic patterns of minima
suitable for synchronization studies.\cite{Grier2003}

%new topic - DM this sentence appeared toward the end of the intro
The model discussed in this paper
resembles a damped driven pendulum 
where external forces
add or remove energy from the system.
A single pendulum is driven by
the external force of gravity to return
to an equilibrium position,
and may be damped to decreasing amplitude
by mechanical friction or air resistance.
Multiple pendula may swing in phase
when coupled through shared vibrations.

External forcing
can cause or regulate synchronization.
For example, 
an electrical pacemaker 
regulates a heart beat 
and 
a pulsed light modifies the
flashing pattern of fireflies. 
%damping
%transition

A single particle oscillator
in a potential well is 
like a skateboarder in a half-pipe or
a child on a swing.
A confined oscillator may 
synchronize its location
to the periodic pattern of the external drive,
moving back and forth in time with
the beat,
or moving between substrate minima.
When 
a constant or 
dc drive is applied,
the particle velocity is modulated by 
the potential energy landscape exerted on the particle.   
Below some threshold  
the dc force is not strong enough to push the particle
across a potential maximum so the average velocity is zero,
a phenomena referred to as pinning.\cite{Reichhardt2017}
Above the pinning threshold,
a particle subject to a constant
drive force  increases its speed at a rate proportional
to the external drive.  
When the applied force varies periodically,
the ac drive can 
cause the particle to hop back and forth across
the landscape minima.
Many synchronized patterns occur,
controlled by the substrate period,
leading to 
mode-locking,
where the average particle velocity
is fixed for a range of dc drive forces.\cite{Reichhardt2015}

In this paper we discuss numerical simulations 
of the synchronized dynamics
of a confined particle driven over
a washboard shaped potential energy landscape.
We describe
our   model  in Sec.~\ref{sec:MD}.
The model
is easy to simulate yet relevant
to condensed matter systems.
We summarize our results 
including the synchronized motion of a single confined particle
driven across a periodic landscape in 
Sec.~\ref{sec:results}. In Sec.~\ref{sec:conclusion}
we describe how our results apply 
to physical systems,
such as optically trapped colloids
and superconducting vortices in the presence of periodic pinning arrays.
We include problems for interested readers 
in Sec.~\ref{sec:problems}.

@@

\section{Conclusion}
\label{sec:conclusion}	
A single particle driven across a periodic potential landscape 
synchronizes its motion 
to environmental and external forces.  
Our simulations reproduce the experiments and simulations presented in 
Juniper {\it et al.} \cite{Juniper2015, Juniper2017}
of 
mode locking in
driven colloids on a
periodic optical landscape.
Colloids are 
relatively easy to 
manipulate and image in experiments,
making ideal proxies 
for systems 
such as cold atoms or electron gases.\cite{Grier2003}
Dynamical mode-locking 
is 
observed in  technologies such as 
 quantum electronic
devices as 
stepped regions in the relation between current and voltage,
where the voltage is the analog of the external driving force
and current is the analog of particle velocity.
%DM FIX FOLLOWING SENTENCE
Known as Shapiro steps, 
these mode-locked or phase-locked currents
%arise 
%due to applied ac voltages
were first
observed in single Josephson junctions\cite{Shapiro1963, Golubov2004}.
%coupled arrays of junctions.\cite{Benz1990}
%DM END FIX
Shapiro steps vary in width depending on the strength of the
applied ac forces,
and are observed in a variety of ac and dc driven systems
displaying
non-Ohmic behavior in voltage-current curves,
including
charge density waves, spin density waves
and superconducting vortices in landscapes 
engineered with periodic patterns of pinning sites.\cite{Reichhardt2000}

>>>>

\section{Introduction} 
Synchronization is a universal phenomena
in which
individual oscillators change frequency due
to external stimuli.\cite{Pikovsky2003}
The
flickering patterns of
candle flames 
mediated by temperature fluctuations,\cite{Okamoto2016}
the vibrations of singing wine glasses interacting 
through sound waves,\cite{Arane2009}
and metronomes vibrating through a supporting platform\cite{Jia2015}
are examples of in-phase coupled oscillations. 
Biological systems benefit from cooperative
synchronization --
birds  coordinate wing flaps
to optimize energy use during flight,\cite{Portugal2014}
frogs alternate croaking patterns,\cite{Aihara2014}
humans clap in time with music,\cite{Tranchant2016}
and 
at a cellular level, 
neurons simultaneously fire in cardiac muscle\cite{MartinHall1999}
and brain tissue.\cite{Singer1999}
%DM ended paragraph with fewer examples, 
%Environmental conditions
%DM add add transition sentence
%
External forcing
can cause or regulate synchronization.
%where external forces
%add or remove energy from the system.
For example, 
an electrical pacemaker 
regulates a heart beat and
a pulsed light modifies the
flashing pattern of fireflies.

The model discussed in this paper
resembles an externally forced pendulum.
%damping
%transition
%%DM - focus more on single particle with environment
A single pendulum 
may swing in time with an %be driven or damped by
driving force
like  %a skateboarder in a half-pipe or
a pushed child on a swing,
and may be damped
by mechanical friction.
%of gravity to return
%to an equilibrium position,
%or  to decreasing amplitude
% or air resistance.
Synchronized phase-locking or mode-locking 
first appeared in the scientific literature
in 1665 with 
Huygens'  experiments on the
motions of synchronized 
pendula in wall-mounted clocks.\cite{Bennett2002}
Clock pendula swing in phase
when coupled through shared vibrations.
A locked-mode is an integer frequency ratio.
In Huygens clocks,
the pendula were observed to swing at the same rate
in the same direction (1:1 mode) or
opposite directions ($-1$:1 mode).
A 2:1 mode occurs when a simple pendulum four
times the length of another swings with twice the period.
%DM transition sentence
Many subsequent 
simulations and laboratory experiments
have revealed 
mode-locked synchronization
in a diverse range of 
oscillation phenomena.

%Single oscillators 
%can demonstrate a variety of synchronized behaviors.
Colloid particles in complex environments are an excellent
model system to explore %these 
the dynamics of mode-locking. %synchronization.
In experiments %DM moved the ``in experiments'' to second sentence
typical colloids are   
plastic spheres suspended in
de-ionized water or silica beads suspended in organic solvent.
Colloids can be trapped
with 
radiation pressure from 
a laser beam.\cite{Ashkin1997} 
A colloid in such a confining region,
an optical trap, oscillates
about a equilibrium point.
A colloid centered in an optical trap is 
uniformly bombarded by photons. 
Off-center colloids 
experience a net force
due to uneven photon collisions across
the particle surface.
Depending on the 
location of the particle in the trap,
the radiation pressure either moves colloids toward the center 
or ejects it from the trap.

% such as synchronized mode-locking can be studied with
%An advantage of colloid studies
While many systems oscillate,
trapped colloids are useful for 
providing controlled 
measurements of the
step-by-step dynamics of particle motion.
Colloids are large 
and move slowly so
%colloids are considered model systems
particle positions 
can be measured in real time 
with an optical camera.\cite{Pertsinidis2001}
Colloids have been
shown to 
provide an ideal system to 
control and study
phase-locked motions
%are useful for
%understanding synchronization
at a single particle level.\cite{Juniper2015,Juniper2017}
%Light is a
%tool for manipulating the colloidal environment,
%which can alter synchronization patterns.
%Synchronization patterns can be altered
%using light to manipulate
%the colloidal environment.
Diffraction gratings can create
more complex light environments, 
such as periodic patterns of minima
suitable for synchronization studies.\cite{Grier2003}
In the following study,
we assume the colloids are confined
to move along a periodic light pattern
with minima and maxima defined by a sine wave.

A colloid driven across a periodic
substrate may 
synchronize its location
to the pattern of the external drive.
%moving back and forth in time with
%the beat, or
%moving between substrate minima.
When 
a constant or 
dc drive is applied,
the particle velocity is modulated by 
the potential energy landscape exerted on the particle.   
Below some threshold  
the dc force is not strong enough to push the particle
across a potential maximum so the average velocity is zero,
a phenomena referred to as pinning.\cite{Reichhardt2017}
Above the pinning threshold,
a dc driven particle
%a particle subject to a constant
%drive force
increases its speed at a rate proportional
to the external drive.  
When the applied force varies periodically,
an ac drive,
%cause
the particle can hop back and forth across
the landscape minima.
Many synchronized patterns occur,
controlled by the substrate pattern and driving force.
When a particle is driven by combined 
ac and dc forces,
mode-locking appears 
where 
the average particle velocity
%of an  driven particle 
is fixed for a range of dc drive forces.\cite{Reichhardt2015}

In this paper we discuss numerical simulations 
of the mode-locked dynamics
of a confined particle driven over
a washboard shaped potential energy landscape.
The numerical model is presented in Sec.~\ref{sec:MD} and 
our results are summarized in
%including the synchronized motion of a single confined particle
%driven across a periodic landscape in 
Sec.~\ref{sec:results}. In Sec.~\ref{sec:conclusion}
we relate  our results 
to a broad variety of physical systems beyond 
optically trapped colloids.
%and superconducting vortices in the presence of periodic pinning arrays.
We include problems for interested readers 
in Sec.~\ref{sec:problems}.

@@

\section{Conclusion}
\label{sec:conclusion}	
A single particle driven across a periodic potential landscape 
synchronizes its motion 
to environmental and external forces.  
Our simulations reproduce the experiments and simulations presented in 
Juniper {\it et al.} \cite{Juniper2015, Juniper2017}
of 
mode locking in
driven colloids on a
periodic optical landscape.
The model
is easy to simulate yet relevant
to a broad range of condensed matter systems.
Colloids are 
relatively easy to 
manipulate and image in experiments,
making ideal proxies 
for systems 
such as cold atoms or electron gases.\cite{Grier2003}
Dynamical mode-locking 
is 
observed in  technologies such as 
 quantum electronic
devices as 
stepped regions in the relation between current and voltage,
where the voltage is the analog of the external driving force
and current is the analog of particle velocity.
%DM FIX FOLLOWING SENTENCE
Known as Shapiro steps, 
these mode-locked or phase-locked currents
%arise 
%due to applied ac voltages
were first
observed in single Josephson junctions\cite{Shapiro1963, Golubov2004}.
%coupled arrays of junctions.\cite{Benz1990}
%DM END FIX
Shapiro steps vary in width depending on the strength of the
applied ac forces,
and are observed in a variety of ac and dc driven systems
displaying
non-Ohmic behavior in voltage-current curves,
including
charge density waves, spin density waves
and superconducting vortices in landscapes 
engineered with periodic patterns of pinning sites.\cite{Reichhardt2000}
